---
title: "ICBC Open Data Demo"
output:
  html_document:
    code_folding: hide
    df_print: paged
---

An example script for accessing open-licensed ICBC data from [https://public.tableau.com/profile/icbc#!/](https://public.tableau.com/profile/icbc#!/)

```{r setup, include=TRUE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

## load R libraries
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidyr)
library(bcdata)

#read in cached data
vehicle_popn <- read_csv(here::here("data/vehicle_popn.csv")) %>% 
    clean_names()

bc_muni <- bcdc_query_geodata('e3c3c580-996a-4668-8bc5-6aa7c7dc4932') %>%
  select(ADMIN_AREA_ABBREVIATION, ADMIN_AREA_GROUP_NAME) %>%
  collect() %>%
  select(ADMIN_AREA_ABBREVIATION, ADMIN_AREA_GROUP_NAME) %>%
  clean_names() %>% 
  sf::st_drop_geometry() %>% 
  mutate(admin_area_abbreviation = gsub(" - District| - City", "", admin_area_abbreviation)) %>% 
  distinct()
```

#### Read in ICBC open-licensed data from the ICBC Public Tableau Dashboard

To download the open data from ICBC, you need to navigate their Tableau interface. This approach has worked to download a CSV file locally:

- Navigate to this page: https://public.tableau.com/profile/icbc#!/vizhome/VehiclePopulation-VehiclePoliciesinForce/2015-2019VehiclePoliciesinForce
- A new page will open with something that looks like a spreadsheet. Scroll to the bottom of the spreadsheet and click on the download icon
- A box will pop up asking you which file format you would like, click on `Data`
- This will open a new window with a `Summary` and `Full Data` tab, click on `Full Data`
- On the `Full Data` tab click on `Download all rows as a text file` and you will be given the option to download the data locally to your machine

```{r get-data, eval=FALSE}
## get open data 

## ICBC Vehicle Population Data provided under ICBC's Open Data Licence
## https://public.tableau.com/profile/icbc#!/vizhome/VehiclePopulationIntroPage/VehiclePopulationData

# Note: this url worked once & then bonked, and manual data button greyed out on the webpage
# maybe hit API limit for the day?
vehicle_popn <- read_csv("https://public.tableau.com/vizql/w/VehiclePopulation-VehiclePoliciesinForce/v/2015-2019VehiclePoliciesinForce/vudcsv/sessions/F921B0B6D90D473BA3D18354807F39E3-0:0/views/5354432173482926677_9740470209173409196?underlying_table_id=Migrated%20Data&underlying_table_caption=Full%20Data") %>% 
    clean_names()
```



#### Count of `Electric` & `GasolineElectric` Vehicles by Year in B.C.



```{r, fig.height=4, fig.width=6, fig.align = 'center'}
#count of electric vehicles by year
vehicle_popn %>% 
  filter(fuel_type %in% c("Electric", "GasolineElectric")) %>% 
  group_by(vehicle_count_year, fuel_type) %>% 
  summarise(total = sum(vehicle_count), .groups = "drop") %>% 
  ggplot(aes(vehicle_count_year, total, colour = fuel_type)) +
  geom_line() +
  geom_point() +
  labs(x = NULL, y = "total vehicle count") +
  theme_minimal() 
```


#### Count of `Electric` + `GasolineElectric` Vehicles by Year for Municipalities

** Plot only includes municipalities with 50 or more `Electric` + `GasolineElectric` Vehicles in 2015


```{r, fig.height=10, fig.width=10, fig.align = 'center'}
##count of electric+gasolinelectric vehicles by year by municipality
ev_by_mun <- vehicle_popn %>%
  filter(fuel_type %in% c("Electric", "GasolineElectric")) %>%
  group_by(vehicle_count_year, municipality) %>%
  summarise(total = sum(vehicle_count), .groups = "drop") %>%
  complete(municipality, vehicle_count_year = 2015:2019,
             fill = list(total = 0)) 


##plot only municipalities where total = or > 100 in 2015
mun_filter <- ev_by_mun %>% 
  filter(total >= 50 & vehicle_count_year == 2015) %>% 
  distinct(municipality) %>% 
  pull()

ev_by_mun_subset <- ev_by_mun %>% 
  filter(municipality %in% mun_filter) %>%  
  left_join(bc_muni, by = c("municipality" = "admin_area_abbreviation"))
  
# generate the plot order based on highest values in most recent year
plot.order <- ev_by_mun_subset %>%
  filter(vehicle_count_year == "2019") %>%
  arrange(desc(total)) %>%
  pull(municipality)


ev_by_mun_subset$municipality = factor(ev_by_mun_subset$municipality, 
                                           levels = plot.order)

ev_by_mun_subset %>% 
  ggplot(aes(vehicle_count_year , total)) +
  geom_line() +
  geom_point() +
  facet_wrap(facets = vars(municipality)) +
  labs(x = NULL, y = "total vehicle count") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5))
```



